Behavior Driven Development in Haskell

### 2012-05-24 paper.tex
writing implementation.

\begin{quote}
\begin{verbatim}
-- | /O(1)/. Is the map empty?
--
-- > Data.Map.null (empty)           == True
-- > Data.Map.null (singleton 1 'a') == False

@@

\begin{quote}
\begin{verbatim}
>>> length []
0
\end{verbatim}
\end{quote}

@@

\begin{quote}
\begin{verbatim}
>>> writeFile "tmpfile" "Hello"
>>> readFile "tmpfile"
"Hello"
\end{verbatim}
\end{quote}

@@

\begin{quote}
\begin{verbatim}
>>> head []
*** Exception: Prelude.head: empty list
\end{verbatim}
\end{quote}

@@

\begin{quote}
\begin{verbatim}
prop> Data.Map.null empty == True
\end{verbatim}
\end{quote}

@@

\begin{quote}
\begin{verbatim}
prop> \xs -> sort xs == sort (sort (xs::[Int]))
\end{verbatim}
\end{quote}

@@

\begin{quote}
\begin{verbatim}
prop> sort xs == sort (sort (xs::[Int]))
\end{verbatim}
\end{quote}

Examples and properties in Haddock annotations are automatically
tested by the {\tt doctest} command. We also provided a library
to do this job. Its implementation will discussed in Section XXX.

@@

\begin{quote}
\begin{verbatim}
TBD: Example
\end{verbatim}
\end{quote}

@@

\begin{quote}
\begin{verbatim}
TBD: Example
Test-Suite spec
  Main-Is: Spec.hs
  Type: exitcode-stdio-1.0
  ...

@@

\begin{quote}
\begin{verbatim}
TBD: Spec.hs comes here
\end{verbatim}
\end{quote}

\begin{quote}
\begin{verbatim}
TBD: doctest.hs comes here
\end{verbatim}
\end{quote}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{doctest}

@@

TBD: how to handle error cases.

TBD: how to integrate QuickCheck.

@@

\begin{quote}
\begin{verbatim}
rev :: [a] -> [a]
rev = undefined
\end{verbatim}
\end{quote}

@@

\begin{quote}
\begin{verbatim}
-- |
-- 'rev' @xs@ returns the elements of @xs@
-- in rev order. @xs@ must be finite.
-- 
-- >>> rev [1,2,3]
-- [3,2,1]
--
-- prop> rev [] == rev []
-- prop> rev (xs ++ ys) == rev ys ++ rev (xs::[Int])
-- prop> rev (rev xs) == (xs::[Int])

@@

\begin{figure*}
\begin{verbatim}
  describe "rev" $ do
    it "return the first element in the last" $ property $
        \xs -> not (null xs) ==> head (rev xs) == last (xs :: [Int])

@@

\begin{quote}
\begin{verbatim}
src/Foo.hs
src/Foo/Bar.hs
src/Baz.hs
\end{verbatim}
\end{quote}

@@

\begin{quote}
\begin{verbatim}
test/FooSpec.hs
test/Foo/BarSpec.hs
test/BazSpec.hs
\end{verbatim}
\end{quote}

@@

\begin{quote}
\begin{verbatim}
import Test.Hspec.Monadic

@@

\begin{quote}
\begin{verbatim}
{-# OPTIONS_GHC -F -pgmF hspec-discover #-}
\end{verbatim}
\end{quote}

>>>>

\begin{quote}
\small
\begin{verbatim}
-- | /O(1)/. Is the map empty?
--
-- > Data.Map.null (empty)           == True
-- > Data.Map.null (singleton 1 'a') == False

@@

\begin{quote}
\small
\begin{verbatim}
>>> length []
0
\end{verbatim}
\end{quote}

@@

\begin{quote}
\small
\begin{verbatim}
>>> writeFile "tmpfile" "Hello"
>>> readFile "tmpfile"
"Hello"
\end{verbatim}
\end{quote}

@@

\begin{quote}
\small
\begin{verbatim}
>>> head []
*** Exception: Prelude.head: empty list
\end{verbatim}
\end{quote}

@@

\begin{quote}
\small
\begin{verbatim}
prop> Data.Map.null empty == True
\end{verbatim}
\end{quote}

@@

\begin{quote}
\small
\begin{verbatim}
prop> \xs -> sort xs == sort (sort (xs::[Int]))
\end{verbatim}
\end{quote}

@@

\begin{quote}
\small
\begin{verbatim}
prop> sort xs == sort (sort (xs::[Int]))
\end{verbatim}
\end{quote}

Examples and properties in Haddock annotations are automatically
tested by the {\tt doctest} command. We also provided a library
to do this job. Its implementation will discussed in Section~\ref{sec:doctest}.

@@

\begin{quote}
\small
\begin{verbatim}
TBD: Example
\end{verbatim}
\end{quote}

@@

\begin{quote}
\small
\begin{verbatim}
TBD: Example
Test-Suite spec
  Main-Is: Spec.hs
  Type: exitcode-stdio-1.0
  ...

@@

\begin{quote}
\small
\begin{verbatim}
TBD: Spec.hs comes here
\end{verbatim}
\end{quote}

\begin{quote}
\small
\begin{verbatim}
TBD: doctest.hs comes here
\end{verbatim}
\end{quote}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{doctest}
\label{sec:doctest}

@@

Since test is done by comparing result string in Haddock notation
and result string from GHCi, exceptions can be handled.

To test properties, we also uses GHCi. 
The `quickCheck'' string is prepended to each property
and GHCi evaluates it.
Success or failures is decided according to result string from GHCi.

If properties contain free variables,
we need to add lambda abstraction prefix before prepending
the `quickCheck'' string.
To check if properties contain free variables, 
we also use GHCi for easy implementation at this moment.
GHCi reports free variables if found:

\begin{quote}
\small
\begin{verbatim}
Prelude Data.List> sort xs == sort (sort (xs::[Int]))
<interactive>:1:6: Not in scope: `xs'
<interactive>:1:24: Not in scope: `xs'
\end{verbatim}
\end{quote}

\noindent In this case, we should prepend the ``\verb|\xs -> |" string
to the property.

@@

\begin{quote}
\small
\begin{verbatim}
rev :: [a] -> [a]
rev = undefined
\end{verbatim}
\end{quote}

@@

\begin{quote}
\small
\begin{verbatim}
-- |
-- 'rev' @xs@ returns the elements of @xs@
-- in rev order. @xs@ must be finite.
-- 
-- >>> rev [1,2,3]
-- [3,2,1]
--
-- prop> rev [] == rev []
-- prop> rev (xs ++ ys) == rev ys ++ rev (xs::[Int])
-- prop> rev (rev xs) == (xs::[Int])

@@

\begin{figure*}
\small
\begin{verbatim}
  describe "rev" $ do
    it "return the first element in the last" $ property $
        \xs -> not (null xs) ==> head (rev xs) == last (xs :: [Int])

@@

\begin{quote}
\small
\begin{verbatim}
src/Foo.hs
src/Foo/Bar.hs
src/Baz.hs
\end{verbatim}
\end{quote}

@@

\begin{quote}
\small
\begin{verbatim}
test/FooSpec.hs
test/Foo/BarSpec.hs
test/BazSpec.hs
\end{verbatim}
\end{quote}

@@

\begin{quote}
\small
\begin{verbatim}
import Test.Hspec.Monadic

@@

\begin{quote}
\small
\begin{verbatim}
{-# OPTIONS_GHC -F -pgmF hspec-discover #-}
\end{verbatim}
\end{quote}
